A game-theory modeling approach to utility and strength of interactions dynamics in biomedical research social networks

### 2017-03-27 casm.tex
Strength of interaction everywhere

\begin{abstract} % abstract
%\parttitle{First part title} %if any
Collaboration has become a cornerstone in biomedical research today.
In contrast to physics which has a long history and experience in
collaborative projects, biology is only recently becoming an evermore
collaborative discipline.
In this article we explore the effect of a collaboration network on the distribution
of players having access to certain amount of resources from other players in the
network and the distribution of the strength of connections among
them. Particularly, we implemented two games played simultaneously:
one for maximizing individual utility based on the iterated Prisoner's
Dilemma; the other, a coordination game for maximizing the connection strength
between players. We are interested in how they affect each other in 
the context of a network of scientific collaboration under the idea that while
researchers are interested in maximizing their utilities, they also know that
it is important to invest in building collaborative relationships.
We tested our simulation on a biomedical research community network in M\'exico and
compared the results with an Erd\"{o}s-Reny\'i, a Watts-Strogatz
small-world and Barab\'asi-Albert topologies. Different topologies
display different global utility and global strength of interaction
distributions. Moreover, the distribution of utility and strength of interaction
in the researchers network is similar to that of 
Barab\'asi-Albert and Watts-Strogatz topologies, respectively. We
believe that utility distribution in the researchers network
suggests that there are socio-cultural mechanisms governing the
network that produce an asymmetric distribution of resources. The
high distribution of strong connections might reflect some sort of
subordination among researchers by which they are morally obliged to
cooperate by the same socio-cultural mechanisms. Finally, the range around
the threshold that regulates the decision to cooperate or defect
according to the agent's historical balance between utility and
strength of collaborative relationships and carrying capacity of the
system is small, suggesting that there is a region in which a phase
transition takes place from a population of cooperators to a
population of defectors. Simulations like ours may help to develop
science policies to promote fair distribution of resources.

@@

After each game, the agent adds up utility ($u$), which is the sum of the
payoffs following the PD matrix. A pair of neighbors will add or substract from the strength of
their interaction ($w_{ji}$) as they coordinate or not, being $w$
also cumulative. We measure global utility and connection strength for the whole
network. Global utility $U$ is the sum of all individual utilities and global
strength of connections or $W$ is the sum of every pair of agents' links
$w$. The strength of interaction can be thought of as some sort of ``trust''.\\

@@

The probability for an agent to cooperate or defect depends on a
number ($\eta$) that refers to a historical balance between average
utility and the average strength of the connections with its
neighbors. It is modeled so because we assume that whatever the result
in utility or strength of connection, as long as one of them
increases, the player will be confident in the strategy followed so
far. In every simulation step, each player will play both games with
every neighbor, but set its decision to cooperate or defect beforehand
by assessing the overall situation of its relationships. This may
sound odd, since one may think that a person decides for each
relationship separately. However in this way players are optimizing
access to resources and the strength of interactions to their
neighbors at the same time, since their state is set by
considering both together.\\

@@

We ran two different experiments. In the first one we simulated different
values of carrying capacity $\nu$. With this experiment we were able
to see how the number of cooperators, utility, strength of connections
among agents and the ratio of shifting state population would change
in the range of the carrying capacity. The states of the agents were
the same at initialization, for all values of the carrying
capacity. Since the model is deterministic, it will return the same
result if run under the same conditions.\\

>>>>

\begin{abstract} % abstract
%\parttitle{First part title} %if any
Collaboration has become a cornerstone in biomedical research today.
In contrast to physics which has a long history and experience in
collaborative projects, biology is only recently becoming an evermore
collaborative discipline.
In this article we explore the effect of a collaboration network on the distribution
of players having access to certain amount of resources from other players in the
network and the distribution of the strength of interactions among
them. Particularly, we implemented two games played simultaneously:
one for maximizing individual utility based on the iterated Prisoner's
Dilemma; the other, a coordination game for maximizing the connection strength
between players. We are interested in how they affect each other in 
the context of a network of scientific collaboration under the idea that while
researchers are interested in maximizing their utilities, they also know that
it is important to invest in building collaborative relationships.
We tested our simulation on a biomedical research community network in M\'exico and
compared the results with an Erd\"{o}s-Reny\'i, a Watts-Strogatz
small-world and Barab\'asi-Albert topologies. Different topologies
display different global utility and global strength of interaction
distributions. Moreover, the distribution of utility and strength of interaction
in the researchers network is similar to that of 
Barab\'asi-Albert and Watts-Strogatz topologies, respectively. We
believe that utility distribution in the researchers network
suggests that there are socio-cultural mechanisms governing the
network that produce an asymmetric distribution of resources. The
high distribution of strong connections might reflect some sort of
subordination among researchers by which they are morally obliged to
cooperate by the same socio-cultural mechanisms. Finally, the range around
the threshold that regulates the decision to cooperate or defect
according to the agent's historical balance between utility and
strength of collaborative relationships and carrying capacity of the
system is small, suggesting that there is a region in which a phase
transition takes place from a population of cooperators to a
population of defectors. Simulations like ours may help to develop
science policies to promote fair distribution of resources.

@@

After each game, the agent adds up utility ($u$), which is the sum of the
payoffs following the PD matrix. A pair of neighbors will add or substract from the strength of
their interaction ($w_{ji}$) as they coordinate or not, being $w$
also cumulative. We measure global utility and strength of interaction for the whole
network. Global utility $U$ is the sum of all individual utilities and global
strength of interaction or $W$ is the sum of every pair of agents' links
$w$. The strength of interaction can be thought of as some sort of ``trust''.\\

@@

The probability for an agent to cooperate or defect depends on a
number ($\eta$) that refers to a historical balance between average
utility and the average strength of the connections with its
neighbors. It is modeled so because we assume that whatever the result
in utility or strength of interaction, as long as one of them
increases, the player will be confident in the strategy followed so
far. In every simulation step, each player will play both games with
every neighbor, but set its decision to cooperate or defect beforehand
by assessing the overall situation of its relationships. This may
sound odd, since one may think that a person decides for each
relationship separately. However in this way players are optimizing
access to resources and the strength of interactions to their
neighbors at the same time, since their state is set by
considering both together.\\

@@

We ran two different experiments. In the first one we simulated different
values of carrying capacity $\nu$. With this experiment we were able
to see how the number of cooperators, utility, strength of interactions
among agents and the ratio of shifting state population would change
in the range of the carrying capacity. The states of the agents were
the same at initialization, for all values of the carrying
capacity. Since the model is deterministic, it will return the same
result if run under the same conditions.\\
